Structured Genealogy Processes

### 2022-01-18 ms.Rnw
sgp update

\begin{figure}
  \resizebox{0.6\linewidth}{!}{
    \begin{tikzpicture}[scale=1.8]
      \usetikzlibrary{arrows}
      \tikzstyle{coordinate}=[inner sep=0pt,outer sep=0pt]
      \tikzstyle{det}=[color=black, thick, >=stealth]
      \tikzstyle{stoch}=[color=black, dashed, thick, >=stealth]
      \node (O) at (0,0) {$\Master_t$};
      \node (X) at (180:2) {$\X_t$};
      \node (H) at (210:2) {$\Hist_t$};
      \node (I) at (240:2) {$\Inv_t$};
      \node (G) at (270:2) {$\Gen_t$};
      \node (P) at (300:2) {$\Prune_t$};
      \node (V) at (330:2) {$\Vis_t$};
      \node (W) at (360:2) {$\Wch_i$};
      \draw [det,->] (O) -- (X);
      \draw [det,->] (O) -- (H);
      \draw [det,->] (H) -- (X);
      \draw [det,->] (O) -- (I);
      \draw [det,->] (O) -- (G);
      \draw [det,->] (O) -- (P);
      \draw [det,->] (O) -- (V);
      \draw [det,->] (O) -- (W);
      \draw [det,->] (V) -- (W);
      \draw [det,->] (G) -- (I) node[midway,below,sloped] {$\mathrm{ext}$};
      \draw [det,->] (G) -- (P) node[midway,below,sloped] {$\mathrm{prune}$};
      \draw [det,->] (P) -- (V) node[midway,below,sloped] {$\mathrm{obs}$};
      \draw [stoch,->] (H) -- (I);
      \draw [stoch,->] (H) -- (G);
      \draw [stoch,->] (H) -- (V);
      \draw [stoch,->] (H) -- (W);
    \end{tikzpicture}
  }
  \caption{
    \label{fig:constellation}
    Relations among the various Markov processes discussed in the paper.
    Deterministic maps are indicated with solid arrows;
    random maps are shown as dashed arrows.
    All the maps shown commute.
    $\X_t$ is the \emph{population process}, a model of the dynamics of some system, which we take as a starting point.
    $\Hist_t$ is the \emph{history process}, which records the full history of $\X_t$.
    $\Inv_t$ is the \emph{inventory process}: at each time $t$, $\Inv_t$ is an inventory of all extant individuals in the population, each of which has a globally unique name.
    $\Gen_t$ is the \emph{genealogy process}, which captures the precise genealogical relationships among all individuals in $\Inv_t$, as well as among any samples that have been taken from the population.
    $\Vis_t$ is the \emph{visible genealogy process}, which is $\Gen_t$ pruned so that only relationships among samples remain.
    Finally $\Wch_i$ is the \emph{embedded chain of the visible genealogy process}, which is $\Vis_{s_i}$, $s_i$ being the time of the $i$-th sample.
    All of these processes can be obtained via deterministic procedures applied to the \emph{master process} $\Master_t$, as described in the text.
  }
\end{figure}

@@

\paragraph{Inventory process, deme process}
First, we define a global set $\Names\coloneqq\N$, which is a collection of infinite unique names. For any Deme $j$ of infective population, we define an inventory $\Inv_t^j(\omega)=\mathrm{inven}\left(\omega\vert_t,j\right)$, a list to document all the extant infected individuals in Deme $j$ at time $t$, by their names, where $\omega\vert_t$ is the sequence of jumps up to time $t$. On top of the global set $\Names$ and the $\mathrm{inven}$ procedure, we can then define six operators: 
\begin{enumerate}[(1)]
  \item $\mathrm{next}(\Names)$ returns the minimum integer in $\Names$, call it $n$, and update the global set $\Names=\Names\setminus\{n\}$;
  \item $\mathrm{initialize}(I)\coloneqq\bigcup_{i=1}^I\{\mathrm{next}(\Names)\}$;
  \item $\mathrm{add}(\sg{I})=\sg{I}\cup\{\mathrm{next}(\Names)\}$;
  \item $\mathrm{drop}(\sg{I},n)=\sg{I}\setminus\{n\}$; and
  \item $\mathrm{movein}(\sg{I},n)=\sg{I}\cup\{n\}$.
\end{enumerate}

@@

Eventually, we can define $\mathrm{inven}(\omega)\coloneqq\{\mathrm{inven}(\omega,j)\}_{j=1}^J$ as a list of inventories, such that $\Inv_t(\omega)\coloneqq\{\Inv_t^j(\omega)\}_{j=1}^J$. We can further define the deme process $\demefuc_t^j(\omega)\coloneqq\frac{\Inv_t^j(\omega)}{\Inv_t(\omega)}$ and $\demefuc_t(\omega)\coloneqq\{\demefuc_t^j(\omega)\}_{j=1}^J$.

@@

Eventually, the current time $t$ and the sequence of nodes, $\Gen=\left(t,\left(\Node{p}_k\right)_{k=0}^{K-1}\right)$, explicitly define a structured genealogy $\Gen$, where $t\in\Time$, $K\in\N$, and node $\Node{p}_k$ are subjected to conditions described in \citet{King2021}. We then use $t(\Gen)$, $K(\Gen)$, and $\Node{p}_k(\Gen)$ to represent the time, the length (\ie the number of nodes), and the $k$-th node of the genealogy $\Gen$, respectively.

\paragraph{Effect of births, death, sampling and transistions}

Births, deaths, samples, and transitions change the topology of a structured genealogy. A death at time $t$ indicates the removal of a leaf in the genealogy and the dismiss of a node. The detailed definition can be found in \citet{King2021} and here we continue to use the $\mathrm{drop}(\Gen)$ to denote the resulting sequence of nodes.

A birth at time $t$ in a structured genealogy is similar to that in a unstructured one \citep{King2021}, while the new member introduced to the queue of sequence is in the same deme as the parent. Let $\Ball{b}$ be the $n$-th black ball, the parent, and there exists a node $\Node{p}\in\Gen$ with $\Ball{b}\in w(\Node{p})=\{\Ball{b},\Ball{b}'\}$ where $\Ball{b}'$ is the other ball held by $\Node{p}$. We then can produce a new node $\Node{p}'=(c,t,\{\Ball{g},\Ball{b}''\},z(\Node{p}))$, where $c=\mathrm{next}(\Names)$, $\Ball{g}=(\green,c)$, and $\Ball{b}''=(\black,c)$. By swapping balls between $\Node{p}$ and $\Node{p}'$, it ends up that $w(\Node{p})=\{\Ball{g},\Ball{b}'\}$ and $w(\Node{p}')=\{\Ball{b},\Ball{b}''\}$. We denote the resulting sequence of nodes as $\mathrm{add}(\seq{\Gen},n)$.

A sample at time $t$ results in the participation of a new node with a new blue ball. The procedure is similar to the previous sampling in \citet{King2021} and the birth mentioned above: the $n$-th black ball held by $\Node{p}$ (\ie $w(\Node{p})=\{\Ball{b},\Ball{b}'\}$) is selected, then a new node $\Node{p}'=(c,t,\{\Ball{g},\Ball{b}''\},z(\Node{p}))$ is generated, where $c=\mathrm{next}(\Names)$, $\Ball{g}=(\green,c)$, and $\Ball{b}''=(\blue,q)$, at the end $\Node{p}$ exchanges $\Ball{b}$ for $\Ball{g}$ with $\Node{p}'$. Here, $q$ is the name of the new blue ball, which is the ordinal number of the transition, $q=|\{\Ball{b}\in w(\Gen)\;|\;\Ball{b}\ \text{is blue}\}|$. Note that, same as that in a birth, $z(\Node{p}')=z(\Node{p})$. We also use the same notation $\mathrm{sample}(\Gen,n)$ as the resulting sequence of nodes.

A transition at time $t$ leads to the inclusion of a new node with a new purple ball. Let $\Ball{b}$ denotes the $n$-th black ball selected to indicate an extant in the population transits from his/her/their current deme to a new one. Assuming that the unique node $\Node{p}\in\Gen$ baring this selected black ball (\ie $\Ball{b}\in w(\Node{p})=\{\Ball{b},\Ball{b}'\}$) is in Deme $i$ (\ie $z(\Node{p})=i$) and it's transiting to Deme $j$ where $j\neq i$. Now we initialize a new node by taking $\Node{p}'=(c,t,\{\Ball{g},\Ball{b}''\},j)$, where $c=\mathrm{next}(\Names)$, $\Ball{g}=(\green,c)$, and $\Ball{b}''=(\purple,q)$. Similar to other events, we swap $\Ball{b}$ for $\Ball{g}$ between nodes $\Node{p}$ and $\Node{p}'$ and insert $\Node{p}'$ at the last position of the node-sequence. Here, $q$ is the name of the new purple ball, which is the ordinal number of the transition, $q=|\{\Ball{b}\in w(\Gen)\;|\;\Ball{b}\ \text{is purple}\}|$. We call this new node $\Node{p}'$ the \emph{Transition Nodes} in the genealogy and denote the resulting sequence of nodes by $\mathrm{transition}(\seq{\Gen},n,i,j)$.

@@

\paragraph{Structured genealogy process}
The structured genealogy evolves over time, so we can proceed to define the \emph{structured genealogy process} $\Gen_t$ now. We define $\Gen_t(\omega) = \left(t,\mathrm{geneal}(\restrict{\omega}{t})\right)$, for the end time $t\in\Time$ and the sequence of jumps $\omega\in\Omega$. The operation $\mathrm{geneal}$ is defined recursively as follows for $\omega\in\mathring\Omega$:

\fontsize{10pt}{12pt}\selectfont
\begin{equation}\label{eq:Gdef}
  \mathrm{geneal}(\omega)\coloneqq
  \begin{cases}
    \left(k,0,\{(\green,k),(\black,k)\},0\right)_{k=0}^{I(\X_0(\omega))}, &\text{if}\ K(\omega)=0;\\
    \mathrm{add}(\mathrm{geneal}(\predec{\omega}),N_{K(\omega)}(\omega)), &\text{if}\ K(\omega)>0\ \text{and}\ U_{K(\omega)}\in\Bset;\\
    \mathrm{drop}(\mathrm{geneal}(\predec{\omega}),N_{K(\omega)}(\omega)), &\text{if}\ K(\omega)>0\ \text{and}\ U_{K(\omega)}\in\Dset;\\
    \mathrm{sample}(\mathrm{geneal}(\predec{\omega}),N_{K(\omega)}(\omega)), &\text{if}\ K(\omega)>0\ \text{and}\ U_{K(\omega)}\in\Sset;\\
    \mathrm{transition}(\mathrm{geneal}(\predec{\omega}),N_{K(\omega)}(\omega),i,j), &\text{if}\ K(\omega)>0\ \text{and}\ U_{K(\omega)}\in\Mset_{ij};\\
    \mathrm{geneal}(\predec{\omega}), &\text{otherwise}.
  \end{cases}
\end{equation}
\normalsize

A deterministic map, $\mathrm{ext}$, can be defined, such that $\Inv_t(\omega)=\mathrm{ext}(\Gen_t(\omega))$ for all $\omega\in\Omega$.

@@

\paragraph{Pruned genealogy process, visible genealogy process}
We divide the procedure of deriving a sampled genealogy from a full genealogy into two steps: \emph{pruning} and \emph{obscuring}. The former one indicates dropping all unsampled individuals (\ie black balls) in a genealogy and remaining the relationship between samples and transition records of samples, and the latter one represents further erasing all deme-transition records of samples and obtaining an ``obscured'' sampled genealogy.

\paragraph{Node color}

\paragraph{Lineage count}

>>>>

\begin{figure}
  \resizebox{0.6\linewidth}{!}{
    \begin{tikzpicture}[scale=1.8]
      \usetikzlibrary{arrows}
      \tikzstyle{coordinate}=[inner sep=0pt,outer sep=0pt]
      \tikzstyle{det}=[color=black, thick, >=stealth]
      \tikzstyle{stoch}=[color=black, dashed, thick, >=stealth]
      \node (A) at (0,0) {$\Master_t$};
      \node (X) at (180:2) {$\X_t$};
      \node (H) at (216:2) {$\Hist_t$};
      \node (I) at (252:2) {$\Inv_t$};
      \node (G) at (288:2) {$\Gen_t$};
      \node (P) at (324:2) {$\Prune_t$};
      \node (V) at (360:2) {$\Vis_t$};
      \node (W) at (342:1.5) {$\Wch_i$};
      \node (O) at (306:1.2) {$\Och_i$};
      \draw [det,->] (A) -- (X);
      \draw [det,->] (A) -- (H);
      \draw [det,->] (H) -- (X);
      \draw [det,->] (A) -- (I);
      \draw [det,->] (A) -- (G);
      \draw [det,->] (A) -- (V);
      \draw [det,->] (A) -- (P);
      \draw [det,->] (A) -- (O);
      \draw [det,->] (P) -- (O);
      \draw [det,->] (A) -- (W);
      \draw [det,->] (V) -- (W);
      \draw [det,->] (G) -- (I) node[midway,below,sloped] {$\mathrm{ext}$};
      \draw [det,->] (G) -- (P) node[midway,below,sloped] {$\mathrm{prune}$};
      \draw [det,->] (P) -- (V) node[midway,below,sloped] {$\mathrm{obs}$};
      \draw [stoch,->] (H) -- (I);
      \draw [stoch,->] (H) -- (G);
      \draw [stoch,->] (H) -- (V);
      \draw [stoch,->] (H) -- (P);
      \draw [stoch,->] (H) -- (O);
      \draw [stoch,->] (H) -- (W);
    \end{tikzpicture}
  }
  \caption{
    \label{fig:constellation}
    Relations among the various Markov processes discussed in the paper.
    Deterministic maps are indicated with solid arrows;
    random maps are shown as dashed arrows.
    All the maps shown commute.
    $\X_t$ is the \emph{population process}, a model of the dynamics of some system, which we take as a starting point.
    $\Hist_t$ is the \emph{history process}, which records the full history of $\X_t$.
    $\Inv_t$ is the \emph{inventory process}: at each time $t$, $\Inv_t$ is an inventory of all extant individuals in the population, each of which has a globally unique name.
    $\Gen_t$ is the \emph{genealogy process}, which captures the precise genealogical relationships among all individuals in $\Inv_t$, as well as among any samples that have been taken from the population.
    $\Vis_t$ is the \emph{visible genealogy process}, which is $\Gen_t$ pruned so that only relationships among samples remain.
    Finally $\Wch_i$ is the \emph{embedded chain of the visible genealogy process}, which is $\Vis_{s_i}$, $s_i$ being the time of the $i$-th sample.
    All of these processes can be obtained via deterministic procedures applied to the \emph{master process} $\Master_t$, as described in the text.
  }
\end{figure}

@@

\paragraph{Inventory process}
First, we define a global set $\Names\coloneqq\N$, which is a collection of infinite unique names. 
For any Deme $j$ of infective population, we define an inventory $\Inv_t^j(\omega)=\mathrm{inven}\left(\omega\vert_t,j\right)$, a list to document all the extant infected individuals in Deme $j$ at time $t$, by their names, where $\omega\vert_t$ is the sequence of jumps up to time $t$. 
On top of the global set $\Names$ and the $\mathrm{inven}$ procedure, we can then define six operators: 
\begin{enumerate}[(1)]
  \item $\mathrm{next}(\Names)$ returns the minimum integer in $\Names$, call it $n$, and update the global set $\Names=\Names\setminus\{n\}$;
  \item $\mathrm{initialize}(I)\coloneqq\bigcup_{i=1}^I\{\mathrm{next}(\Names)\}$;
  \item $\mathrm{add}(\sg{I})=\sg{I}\cup\{\mathrm{next}(\Names)\}$;
  \item $\mathrm{drop}(\sg{I},n)=\sg{I}\setminus\{n\}$; and
  \item $\mathrm{movein}(\sg{I},n)=\sg{I}\cup\{n\}$.
\end{enumerate}

@@

Eventually, we can define $\mathrm{inven}(\omega)\coloneqq\{\mathrm{inven}(\omega,j)\}_{j=1}^J$ as a list of inventories, such that $\Inv_t(\omega)\coloneqq\{\Inv_t^j(\omega)\}_{j=1}^J$.

@@

Eventually, the current time $t$ and the sequence of nodes, $\Gen=\left(t,\left(\Node{p}_k\right)_{k=0}^{K-1}\right)$, explicitly define a structured genealogy $\Gen$, where $t\in\Time$, $K\in\N$, and node $\Node{p}_k$ are subjected to conditions described in \citet{King2021}. 
We then use $t(\Gen)$, $K(\Gen)$, and $\Node{p}_k(\Gen)$ to represent the time, the length (\ie the number of nodes), and the $k$-th node of the genealogy $\Gen$, respectively. 
We define $\seq{\Gen}$ as the node sequence of genealogy $\Gen$, where $\Node{p}\in\Gen$ when $\Node{p}$ is one node in $\seq{\Gen}$.

\paragraph{Effect of births, death, sampling and transistions}
Births, deaths, samples, and transitions change the topology of a structured genealogy. 
A death at time $t$ indicates the removal of a leaf in the genealogy and the dismiss of a node. The detailed definition can be found in \citet{King2021}.

A birth at time $t$ in a structured genealogy is similar to that in a unstructured one \citep{King2021}: a black ball is selected as the parent and a new node introduced as the child, who is in the same deme as the parent. Let $\Ball{b}$ be randomly selected black ball, named \emph{n}, which is the parent, and there exists a node $\Node{p}\in\Gen$ holding $\Ball{b}\in w(\Node{p})=\{\Ball{b},\Ball{b}'\}$ where $\Ball{b}'$ is the other ball held by $\Node{p}$. 
Here, we assume the parent is from Deme $i$ (\ie $n\in\Inv_i$) and the child is born into Deme $j$. 
We then can produce a new node $\Node{p}'=(c,t,\{\Ball{g},\Ball{b}''\},i)$, where $c=\mathrm{next}(\Names)$, $\Ball{g}=(\green,c)$, $\Ball{b}''=(\black,c)$, and update $\Inv_j=\Inv_j\cup\{c\}$. 
By swapping balls between $\Node{p}$ and $\Node{p}'$, it ends up that $w(\Node{p})=\{\Ball{g},\Ball{b}'\}$ and $w(\Node{p}')=\{\Ball{b},\Ball{b}''\}$. 
We denote the resulting sequence of nodes as $\mathrm{add}(\seq{\Gen},n)$.

A death at time $t$ is defined in the same way as that in \citet{King2021} and we continue to use the notation $\mathrm{drop}(\seq{\Gen},n)$ to denote the resulting sequence of nodes for a death, where an individual with name \emph{n} is selected randomly to be removed from the population.

A sample at time $t$ results in the participation of a new node with a new blue ball. 
The procedure is similar to the previous sampling in \citet{King2021} and the birth mentioned above: a black ball, named \emph{n}, where $n\in\Inv_j$, and held by $\Node{p}$ (\ie $w(\Node{p})=\{\Ball{b},\Ball{b}'\}$), is selected at random. 
Then a new node $\Node{p}'=(c,t,\{\Ball{g},\Ball{b}''\},j)$ is generated, where $c=\mathrm{next}(\Names)$, $\Ball{g}=(\green,c)$, and $\Ball{b}''=(\blue,q)$. 
At the end, $\Node{p}$ exchanges $\Ball{b}$ for $\Ball{g}$ with $\Node{p}'$. 
Here, $q$ is the name of the new blue ball, which is the ordinal number of the transition, $q=|\{\Ball{b}\in w(\Gen)\;|\;\Ball{b}\ \text{is blue}\}|$. 
We use the notation $\mathrm{sample}(\seq{\Gen},n)$ as the resulting sequence of nodes.

A transition at time $t$ leads to the inclusion of a new node with a new purple ball. 
Let $\Ball{b}$ denote the
black ball with name \emph{n} and selected to indicate an extant in the population transits from his/her current
deme to a new one. 
Assuming this selected individual is from Deme $i$ (\ie $i\in\Inv_i$) and transiting to Deme $j$ ($j\neq i$), we then initialize a new node by taking $\Node{p}'=(c,t,\{\Ball{g},\Ball{b}''\},i)$, where $c=\mathrm{next}(\Names)$, $\Ball{g}=(\green,c)$, and $\Ball{b}''=(\purple,q)$. 
Note that, we update $\Inv_i=\Inv_i\setminus\{n\}$ and $\Inv_j=\Inv_j\cup\{n\}$. 
Similar to other events, we therefore swap $\Ball{b}$ for $\Ball{g}$ between nodes $\Node{p}$ and $\Node{p}'$ and insert $\Node{p}'$ at the last position of the node-sequence. 
Here, \emph{q} is the name of the new purple ball, which is the ordinal number of the transition,  $q=|\{\Ball{b}\in w(\Gen)\;|\;\Ball{b}\ \text{is purple}\}|$. 
We call this new node $\Node{p}'$ the \emph{Transition Nodes} in the genealogy and denote the resulting sequence of nodes by $\mathrm{transition}(\seq{\Gen},n)$.

@@

\paragraph{Structured genealogy process}
The structured genealogy evolves over time, so we can proceed to define the \emph{structured genealogy process} $\Gen_t$ now. 
We define $\Gen_t(\omega) = \left(t,\mathrm{geneal}(\restrict{\omega}{t})\right)$, for the end time $t\in\Time$ and the sequence of jumps $\omega\in\Omega$. 
We first initialize the global set of names by letting $\Names=\N\setminus\{1,2,\dots,I(\X_0(\omega))\}$, then the operation $\mathrm{geneal}$ is defined recursively as follows for $\omega\in\mathring\Omega$:

\fontsize{10pt}{12pt}\selectfont
\begin{equation}\label{eq:Gdef}
  \mathrm{geneal}(\omega)\coloneqq
  \begin{cases}
    \left(k,0,\{(\green,k),(\black,k)\},0\right)_{k=0}^{I(\X_0(\omega))}, &\text{if}\ K(\omega)=0;\\
    \mathrm{add}(\mathrm{geneal}(\predec{\omega}),N_{K(\omega)}(\omega)), &\text{if}\ K(\omega)>0\ \text{and}\ U_{K(\omega)}\in\Bset;\\
    \mathrm{drop}(\mathrm{geneal}(\predec{\omega}),N_{K(\omega)}(\omega)), &\text{if}\ K(\omega)>0\ \text{and}\ U_{K(\omega)}\in\Dset;\\
    \mathrm{sample}(\mathrm{geneal}(\predec{\omega}),N_{K(\omega)}(\omega)), &\text{if}\ K(\omega)>0\ \text{and}\ U_{K(\omega)}\in\Sset;\\
    \mathrm{transition}(\mathrm{geneal}(\predec{\omega}),N_{K(\omega)}(\omega)), &\text{if}\ K(\omega)>0\ \text{and}\ U_{K(\omega)}\in\Mset;\\
    \mathrm{geneal}(\predec{\omega}), &\text{otherwise};
  \end{cases}
\end{equation}
\normalsize

where $\Bset\coloneqq\bigcup_{i,j}\Bset_{ij}$, $\Dset\coloneqq\bigcup_j\Dset_j$, and $\Mset\coloneqq\bigcup_{i,j}\Mset_{ij}$. 
A deterministic map, $\mathrm{ext}$, can be defined, such that $\Inv_t(\omega)=\mathrm{ext}(\Gen_t(\omega))$ for all $\omega\in\Omega$.

@@

\paragraph{Pruned genealogy process}
The sampled genealogy contains only relations among genomic samples. 
In other words, information in a full genealogy about unsampled individuals and deme transitions are unobserved in a sampled genealogy, which we call it a visible genealogy. 
We divide the procedure of deriving a visible genealogy from a full genealogy into two separate steps: \emph{pruning} and \emph{obscuring}. 
The former drops all unsampled individuals (\ie black balls) in a genealogy and remains the relationship between samples and deme-transition records of samples, and the latter further erases all deme-transition records of samples and obtains an obscured genealogy.

Given a genealogy $\Gen$, we define $\Inv=\mathrm{ext}(\Gen)$.
%% Then. by definition, for each $c\in\Inv$, $(\black,c)\in w(\Gen)$.
Let $\mathrm{prune}(\Gen)$ be the pruned genealogy derived by applying iterative $\mathrm{drop}$ operation to $\Gen$ for each name $c\in\Inv$.
Specifically, suppose $\Inv^j=\{n_0^j,\dots,n_{m_j-1}^j\}$ is the inventory for Deme $j$ and $\Inv=\{\Inv^j\}_{j=1}^{J}$, where $J$ is the total number of demes. 
Then we can define the set of all names in $\Inv$ as $\mathcal{N}\coloneqq\bigcup_{j=1}^J \Inv^j=\{c_0,\dots,c_{m-1}\}$, where $m=\sum_j m_j$.
Let $\Node{P}_0=\seq\Gen$ and $\Node{P}_{k}=\mathrm{drop}(\Node{P}_{k-1},c_{k-1})$, for $k=1,\dots,m$.
Then $\mathrm{prune}(\Gen)\coloneqq(t(\Gen),\Node{P}_{m})$.
\Cref{fig:prune} illustrates the pruning procedure.

For $\omega\in\Omega$, we define the \emph{pruned genealogy process},
\begin{equation}
  \Prune_t(\omega)\coloneqq\textrm{prune}(\Gen_t(\omega)).
\end{equation}
Note that $\Prune_t$ so defined is a itself a genealogy.
\Cref{fig:prune}B depicts one example of $\Prune_t$ both graphically (as a tree) and diagrammatically (as a node-sequence).

\paragraph{Obscured genealogy process}
We first define an operation $\mathrm{reverse}$ on the sequence of nodes of a genealogy as follows: 
\begin{inparaenum}[(1)]
  \item a random purple ball $\Ball{b}$ with name $q$ is selected such that there exists a node $\Node{p}$, where $\Ball{b}\in w(\Node{p})=\{\Ball{b}, \Ball{b}'\}$ and $\Ball{b}'$ is the other ball held by $\Node{p}$;
  \item there also exists a node $\Node{p}'$ such that $\Ball{g}=\left(\green,n(\Node{p})\right)\in w(\Node{p}')=\{\Ball{g},\Ball{b}''\}$ where $\Ball{b}''$ is the other ball held by $\Node{p}'$;
  \item swapping balls between $\Node{p}$ and $\Node{p}'$ results in $w(\Node{p}')=\{\Ball{b}',\Ball{b}''\}$ and $w(\Node{p})=\{\Ball{g},\Ball{b}\}$;
  \item we delete $\Node{p}$ from the node sequence.
\end{inparaenum}
The resulting sequence of nodes is denoted as $\mathrm{reverse}(\seq{\Gen},q)$.

Suppose for a genealogy $\Gen$, the set of the serial number of the purple balls is $\{0,1,\dots,q\}$ and the list of inventories is $\Inv=\{\Inv^j\}_{j=1}^J$. $\mathrm{obs}(\Gen)$ applies the following steps on $\Gen$: 
\begin{inparaenum}[(a)]
  \item let $\Node{P}_0=\seq{\Gen}$ and $\Node{P}_{k}=\mathrm{reverse}(\Node{P}_{k-1},k)$, for $k=1,\dots,q$; and
  \item for $\Node{P}_q$, we update $\Inv=\bigcup_{j}\Inv^j$ and for each node $\Node{p}\in\seq{\Gen}$, $z(\Node{p})=0$.
\end{inparaenum}

For $\omega\in\Omega$, we define the \emph{obscured genealogy process},
\begin{equation}
  \Obscure_t(\omega)\coloneqq\textrm{obs}(\Gen_t(\omega)).
\end{equation}
Note that $\Obscure_t$ so defined is a itself a genealogy.
\Cref{fig:obs}B depicts one example of $\Obscure_t$ both graphically (as a tree) and diagrammatically (as a node-sequence).

\paragraph{Visible genealogy process}

\paragraph{Node color}
There are at most three kinds of nodes in a visible genealogy, distinguished by the contents of their pockets:
\begin{inparaenum}[(a)]
\item \emph{green nodes}, which have two green balls in their pockets;
\item \emph{blue nodes}, which have one blue and one green ball;
\item \emph{red nodes}, which have one red and one blue ball.
\end{inparaenum}
This distinction allows a compact graphical representation of the visible genealogy (\cref{fig:prune}C) and proves useful in deriving our main results as well.
Green nodes correspond to \emph{coalescence points} (branch points) in the visible genealogy;
red nodes, to leaves;
blue nodes correspond to direct-descent events, i.e., samples which are directly ancestral to other samples.
Thus, if $\Vis$ is a visible genealogy,
$\Eset(\Vis)=\Cset(\Vis)\cup\Qset(\Vis)\cup\Lset(\Vis)$, where
$\Cset(\Vis)$ is the set of times of the green nodes,
$\Qset(\Vis)$ contains the times of the blue nodes, and
$\Lset(\Vis)$ holds the times of the red nodes.
Moreover, $\Tset(\Vis)=\Qset(\Vis)\cup\Lset(\Vis)$ is the set of sample times and $\Masterset(\Vis)=\Qset(\Vis)\cup\Cset(\Vis)$ comprises the \emph{attachment times},
i.e., those points at which a sampled lineage attaches to the visible genealogy subtended by earlier samples.

\paragraph{Deme proportion and lineage count}

We can further define the deme proportion function $\demefuc(t,\omega)\coloneqq\{\demefuc^j(t,\omega)\}_{j=1}^J$, where $\demefuc^j(t,\omega)\coloneqq\frac{\vert\Inv_t^j(\omega)\vert}{\vert\Inv_t(\omega)\vert}$.

\QYL{Lineage count function may not be a \cadlag}
